Loading and preprocessing the data
=======================================

```{r}
library(knitr)
mydata<-read.csv("activity.csv",head=TRUE,sep=",")
nrow(mydata)
str(mydata)
data<-na.omit(mydata)
nrow(data)
```

What is mean total number of steps taken per day?
==================================================
```{r}
library(dplyr)
activity<-data %>% group_by(date) %>% summarize(totalsteps=sum(steps))

hist(activity$totalsteps,xlab="Number steps per each day",ylab="Number of days",col="pink",main="Graphical Representation Of Total Steps In a Day")

datamean<-mean(activity$totalsteps)
datamedian<-median(activity$totalsteps)
 
print(sprintf("Mean of total number of steps taken per day is: %f",datamean))
print(sprintf("Meadian of total number of steps taken per day is: %f",datamedian))
```

What is the average daily activity pattern ?
================================================

```{r}
dtq2<-data %>% group_by(interval) %>% summarize(average=mean(steps))

plot(dtq2$interval,dtq2$average,xlab="Interval",ylab="Average number of steps",type="l",col="blue",main="Average Daily Activity Pattern")

maxvalue<-dtq2$interval[dtq2$average==max(dtq2$average)]

print(sprintf("5-min time Interval with maximum average steps taken per day: %i ",maxvalue))
```

Imputing Missing Values
========================

```{r}
missingrows<- (nrow(mydata)-nrow(data))
print(sprintf("Total number of missing values in the dataset(ie.,NA's): %i", missingrows))
```

Strategy
===========


Replace NA step values for corresponding intervals with mean interval values calculated above
===========================

```{r}
newdata<-mydata
imdt = mean(mydata$steps,na.rm=T)
newdata[,"steps"][is.na(newdata$steps)]=imdt
aggdata <- aggregate(x = newdata[c("steps")], FUN = sum,by = list(Group.date = newdata$date),na.rm=TRUE)
aggdata
```

Plot Histogram of total steps with imputed data
================================================
```{r}
hist(aggdata$steps,xlab="Total steps per day",ylab="Number of Days",main="Total Steps Per Day",col="green")
```

Calculate and report the mean and median total number of steps taken per day
====================================
```{r}
 ty<-newdata %>% group_by(date) %>% summarize(totalsteps=sum(steps))
meansteps <- mean(ty$totalsteps)
mediansteps <- median(ty$totalsteps)

print(sprintf("Mean total steps taken per day: %f ", meansteps))
print(sprintf("Median total steps taken per day: %f ", mediansteps))
```

Are there differences in activity patterns between weekdays and weekends?
======================================================

```{r}
mydata = read.csv("activity.csv")
wdata<-mydata
wdata$day=weekdays(as.Date(wdata$date))
wdata$weekday[wdata$day %in% c("Saturday","Sunday")] = "Weekend"
wdata$weekday[!(wdata$day %in% c("Saturday","Sunday"))] = "Weekday"

table(wdata$weekday,wdata$day)

```

Graphical Representation
========================
 
```{r}  
aggdata <- aggregate(x = wdata[c("steps")], FUN = sum,by = list(Group.interval = wdata$interval, weekday=wdata$weekday),na.rm=TRUE)
library(lattice)
xyplot(steps~Group.interval|weekday, data=aggdata, type="l",col="blue",xlab="Interval",ylab="Average Number of Steps")

```

